---
title: "MAIHDA_population_level"
author: "Victoria Fisher"
date: "2025-01-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(pacman)

p_load(tidyverse,
       tidycensus,
       ggplot2,
       gtsummary,
       DescTools,
       datawizard,
       excessmort,
       lubridate,
       readxl,
       haven,
       MASS,
       lme4,
       lmerTest,
       nlme,
       glmmTMB,
       brms,
       merTools,
       bbmle,
       zctaCrosswalk,
       performance,
       purrr,
       ggpubr,
       foreign,
       DHARMa,
       sf,
       rms,
       mlogit,
       effects,
       splines,
       car,
       Publish,
       msme,
       plm,
       broom,
       forestplot,
       sjPlot,
       sjlabelled,
       sjmisc,
       clusterSim,
       geepack,
       geeM,
       cardx,
       insight,
       parameters,
       ggeffects,
       emmeans)

options(tigris_use_cache = FALSE)

```

#Loading datasets

```{r}

#Covid mortality

deaths_20_23 <- read_delim('COVID Mortality Statistics 2020 through 2023.txt', 
    delim = "\t", escape_double = FALSE, 
    col_types = cols(Notes = col_skip()), 
    trim_ws = TRUE)

colnames(deaths_20_23)[2] <- "GEOID"
#problems(deaths_20_23)

#ACS

v2020 <- load_variables(2020, "acs5")

vars <- c(county_male_pop = "B01001_002",
          county_white_pop = "B02001_002",
          median_age_total = "B01002_001",
          median_income_12_mo = "B06011_001"
)

years <- 2020:2023 #pulls American Community Survey data from 2019 - 2021
names(years) <- years
vars_county <- purrr::map_dfr(years, ~{
  get_acs(
    geography = "county", 
    variables = vars, 
    year = .x,
    geometry = FALSE
    ) 
}, .id = "year")

vars_county <- vars_county %>%
  pivot_wider(id_cols = c("year", "GEOID"),
              names_from = c("variable"),
              values_from = "estimate")


#Total pop and geography
years <- 2020:2023 #pulls American Community Survey data from 2020 - 2023
names(years) <- years
pop_county <- purrr::map_dfr(years, ~{
  get_acs(
    geography = "county", 
    variables = "B01001_001", 
    year = .x,
    geometry = TRUE
    ) 
}, .id = "year") %>% tigris::shift_geometry()

colnames(pop_county)[5] = "total_pop"
pop_county <- pop_county %>% dplyr::select(!c(4,6))

#Merge two datasets

vars_county = merge(pop_county, vars_county, by = c("year", "GEOID"))

#Percentages
colnames(vars_county)
percent_cols <- names(vars_county[, c(5, 7)])

vars_county <- vars_county %>%
  mutate(across(all_of(percent_cols), ~ (. / total_pop) * 100, .names = "P_{.col}"))

vars_county_avg <- vars_county %>%
  group_by(GEOID) %>%
  summarise(across(where(is.numeric), ~mean(.x)))

#Merge with death dataset

df_mort <- merge(deaths_20_23, vars_county_avg, by = "GEOID")

colnames(df_mort)

```

#Strata

```{r}

summary(is.na(df_mort)) #median income

#Imputing NA median income

summary(is.na(df_mort$median_income_12_mo))

med_income_na <- filter(df_mort, is.na(df_mort$median_income_12_mo))
#25-34 range

df_mort$median_income_12_mo <- ifelse(is.na(df_mort$median_income_12_mo), 29500, df_mort$median_income_12_mo)


#Strata: age, sex, income, race

#Income


df_mort <- df_mort %>%
  mutate(
    # Calculate Q1, Q3, and IQR for 'value' (you can replace 'value' with any column)
    Q1 = quantile(median_income_12_mo, 0.25, na.rm = TRUE),
    Q3 = quantile(median_income_12_mo, 0.75, na.rm = TRUE),
    
    # Create categories based on IQR
    strata_income_cat = case_when(
      median_income_12_mo < Q1 ~ "low median income",                # Below the first quartile
      median_income_12_mo >= Q1 & median_income_12_mo < median(median_income_12_mo, na.rm = TRUE) ~ "med-low median income", # Between Q1 and Q3
      median_income_12_mo >= median(median_income_12_mo, na.rm = TRUE) & median_income_12_mo <= Q3 ~ "med-high median income", # Between Q1 and Q3
      median_income_12_mo > Q3 ~ "high median income",               # Above the third quartile
      TRUE ~ "low median income"                   # NA values indicate low median individual income levels by bracket
    )
  )

table(df_mort$strata_income_cat)


#Age
summary(df_mort$median_age_total)
hist(df_mort$median_age_total)

df_mort <- df_mort %>%
  mutate(
    # Calculate Q1, Q3, and IQR for 'value' (you can replace 'value' with any column)
    Q1 = quantile(median_age_total, 0.25, na.rm = TRUE),
    Q3 = quantile(median_age_total, 0.75, na.rm = TRUE),
    
    # Create categories based on IQR
    strata_age_cat = case_when(
      median_age_total < Q1 ~ "1_age",                # Below the first quartile
      median_age_total >= Q1 & median_age_total <= median(median_age_total) ~ "2_age", # Between Q1 and Q3
      median_age_total > median(median_age_total) & median_age_total <= Q3 ~ "3_age", # Between Q1 and Q3
      median_age_total > Q3 ~ "4_age",               # Above the third quartile
      TRUE ~ "Unknown"                   # In case there are NA values
    )
  )

table(df_mort$strata_age_cat)

#Sex

df_mort$strata_sex_cat <- ifelse(df_mort$P_county_male_pop > 50, "Maj_Male", "Maj_Female")

table(df_mort$strata_sex_cat)

#Race (without ethnicity)

df_mort$P_non_white_only_pop <- 100 - df_mort$P_county_white_pop

df_mort$strata_race_cat <- cut(df_mort$P_non_white_only_pop, breaks = quantile(df_mort$P_non_white_only_pop, probs = 0:2 / 2), include.lowest = TRUE, labels = c("below median", "above median"))

by(df_mort$P_non_white_only_pop, df_mort$strata_race_cat, summary)

ggplot(df_mort, aes(P_non_white_only_pop)) +
  geom_histogram() +
  facet_grid(df_mort$strata_race_cat)

#Composite strata

df_mort$strata <- as.factor(paste(df_mort$strata_sex_cat, ",",
                        df_mort$strata_age_cat, ",",
                        df_mort$strata_race_cat, ",",
                        df_mort$strata_income_cat))

length(unique(df_mort$strata)) #64

```

#Mortality

```{r}

df_mort$covid_deaths <- as.numeric(df_mort$Deaths) #NAs introduced

df_mort$Population <- as.numeric(df_mort$Population)

df_mort$Population <- ifelse(is.na(df_mort$Population), df_mort$total_pop, df_mort$Population)
summary(df_mort$Population)

#Imputing suppressed deaths
df_mort <- df_mort %>%
  mutate(deaths_imputed = ifelse(is.na(covid_deaths), sample(1:4, sum(is.na(covid_deaths)), replace = TRUE), covid_deaths))

summary(df_mort$deaths_imputed)

#Crude rate

df_mort$covid_mortality_rate_100k <- (df_mort$deaths_imputed / df_mort$Population) * 100000
summary(df_mort$covid_mortality_rate_100k) #Almost normally distributed
boxplot(df_mort$covid_mortality_rate_100k)

#Separating county and state
df_mort <- separate_wider_delim(df_mort, cols = `Residence County`, delim = ",", names = c("County", "State"))


#Outliers 95%

outliers <- as.data.frame(check_outliers(df_mort$covid_mortality_rate_100k))

df_mort_95 <- df_mort[!outliers$Outlier, ]

hist(df_mort_95$covid_mortality_rate_100k) # ostensibly normally distributed

#As factors

df_mort_95$strata_age_cat <- as.factor(df_mort_95$strata_age_cat)
df_mort_95$strata_income_cat <- as.factor(df_mort_95$strata_income_cat)
df_mort_95$strata_sex_cat <- as.factor(df_mort_95$strata_sex_cat)
df_mort_95$strata_race_cat <- as.factor(df_mort_95$strata_race_cat)

```


```{r}

#Checking cell counts of strata is sufficient for analysis

strata_df <- df_mort_95 %>%
  group_by(strata) %>%
  mutate(strataN = n())

total_rows <- length(strata_df$strataN)
rows_with_20_or_more <- sum(strata_df$strataN >= 20) 
percentage <- (rows_with_20_or_more / total_rows) * 100

print(percentage) #95

strata_sum <- df_mort_95 %>%
  group_by(strata) %>%
  summarise(across(where(is.numeric), ~ mean(., na.rm = TRUE)))

ggplot(strata_df, aes(reorder(strata, covid_mortality_rate_100k, FUN = median), covid_mortality_rate_100k))+
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(strata_df, aes(reorder(strata, strataN, FUN = median)))+
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  xlab("Strata")

#Descriptive statistics

colnames(df_mort_95)

table1 <- df_mort_95 %>% dplyr::select(c("covid_mortality_rate_100k", "median_income_12_mo", "median_age_total", "P_county_male_pop", "P_non_white_only_pop"))

Desc(table1)

#Graphs of within-group variance

ggplot(df_mort_95, aes(strata_sex_cat, P_county_male_pop))+
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

regression_table2b <- df_mort_95 %>% dplyr::select(c("strata_age_cat",  "strata_race_cat", "strata_sex_cat", "strata_income_cat", "median_income_12_mo", "median_age_total", "P_county_male_pop", "P_non_white_only_pop"))


```

#Model selection    

```{r}

lmer1 <- lmer(covid_mortality_rate_100k ~ 1 + (1 | strata), df_mort_95)
summary(lmer1)
check_model(lmer1)
AIC(lmer1) #30847.34

glmm1 <- glmmTMB(covid_mortality_rate_100k ~ 1 + (1 | strata), df_mort_95, family = "gaussian")
summary(glmm1)
check_model(glmm1)
AIC(glmm1) #30851.03

brms1 <- brm(covid_mortality_rate_100k ~ 1 + (1 | strata), df_mort_95, family = "gaussian")
summary(brms1)
check_model(brms1)
loo(brms1) #looic 30694.8


compare_performance(lmer1, glmm1, rank = TRUE) #glmm1 is preferred, but LMER is similar. LMER for manuscript

```



#GLMM (for manuscript)

```{r}
#Unadjusted

glmm1 <- glmmTMB(covid_mortality_rate_100k ~ 1 + (1 | strata_sex_cat:strata_age_cat:strata_income_cat:strata_race_cat), df_mort_95, family = "gaussian")
summary(glmm1)
confint(glmm1)
AIC(glmm1) #30735.53
icc(glmm1)$ICC_adjusted #0.3333385
check_model(glmm1)

537.2 / (537.2 + 1074.3) #0.333354

plot(ggpredict(glmm1, terms = c("strata_age_cat", "strata_race_cat", "strata_sex_cat",  "strata_income_cat"),
               type = "random"))

#Adjusted with continuous strata

glmm2 <- glmmTMB(covid_mortality_rate_100k ~ standardise(median_age_total) + standardise(median_income_12_mo) + standardise(P_county_male_pop) + standardise(P_non_white_only_pop) + (1 | strata_sex_cat:strata_age_cat:strata_income_cat:strata_race_cat), df_mort_95, family = "gaussian")
summary(glmm2)
confint(glmm2)
AIC(glmm2) #30535.35
icc(glmm2)$ICC_adjusted #0.1199105
check_model(glmm2)

140.2 / (140.2 + 1029.2) #0.1198905

#Adjusted with categorical strata

glmm3 <- glmmTMB(covid_mortality_rate_100k ~ strata_sex_cat + strata_age_cat + strata_income_cat + strata_race_cat + (1 | strata_sex_cat:strata_age_cat:strata_income_cat:strata_race_cat), df_mort_95, family = "gaussian")
summary(glmm3)
confint(glmm3)
AIC(glmm3) #30623.46
icc(glmm3)$ICC_adjusted #0.03507287
check_model(glmm3)

39.15 / (39.15 + 1077.10 ) * 100 #3.507279

#K-fold cross validation using MAE

model_formula <- covid_mortality_rate_100k ~ strata_sex_cat + strata_age_cat + strata_income_cat + strata_race_cat + (1 | strata_sex_cat:strata_age_cat:strata_income_cat:strata_race_cat)

# k-fold function
k_fold_cv_glmmTMB <- function(data, k = 20, formula) {
  
  # Vector to store RMSE or other metrics for each fold
  rmse_values <- numeric(k)
  
  # Randomly split data into k folds
  set.seed(42)  # For reproducibility
  folds <- sample(1:k, nrow(data), replace = TRUE)
  
  # Loop through each fold
  for (i in 1:k) {
    
    # Create training and test sets based on the fold
    test_indices <- which(folds == i)
    train_indices <- setdiff(1:nrow(data), test_indices)
    
    # Create training and test datasets
    train_data <- data[train_indices, ]
    test_data <- data[test_indices, ]
    
    # Fit the model using glmmTMB on the training data
    model <- glmmTMB(formula, data = train_data)
    
    # Predict on the test set
    predictions <- predict(model, newdata = test_data, type = "response")
    
    # Calculate MAE for the test set
     rmse_values[i] <- sqrt(mean((predictions - test_data$covid_mortality_rate_100k)^2))
  }
  
  # Return the average MAE across all folds
  mean_rmse <- mean(rmse_values)
  return(mean_rmse)
}


# Run 10-fold cross-validation for the multilevel model
cv_result <- k_fold_cv_glmmTMB(df_mort_95, k = 20, formula = model_formula)

# View the average RMSE result
print(cv_result) #33

#Single-line model

glmm4 <- glmmTMB(covid_mortality_rate_100k ~ strata_sex_cat:strata_age_cat:strata_income_cat:strata_race_cat, df_mort_95, family = "gaussian")
summary(glmm4)
AIC(glmm4) #30597
check_model(glmm4)
performance::r2(glmm4) #.352
confint(glmm4)

compare_performance(glmm1, glmm2, glmm3, glmm4)
anova(glmm1, glmm2, glmm3, glmm4)

#Sensitivity analysis Model 5

glmm5 <- glmmTMB(covid_mortality_rate_100k ~ standardise(median_age_total)*standardise(median_income_12_mo)*standardise(P_county_male_pop)*standardise(P_non_white_only_pop) + (1 | strata_sex_cat:strata_age_cat:strata_income_cat:strata_race_cat), df_mort_95, family = "gaussian")
summary(glmm5)
AIC(glmm5) #30503.24
confint(glmm5)
check_model(glmm5)

#K-fold cross validation using MAE

model_formula <- covid_mortality_rate_100k ~ standardise(median_age_total)*standardise(median_income_12_mo)*standardise(P_county_male_pop)*standardise(P_non_white_only_pop) + (1 | strata_sex_cat:strata_age_cat:strata_income_cat:strata_race_cat)

# k-fold function
k_fold_cv_glmmTMB <- function(data, k = 20, formula) {
  
  # Vector to store RMSE or other metrics for each fold
  rmse_values <- numeric(k)
  
  # Randomly split data into k folds
  set.seed(42)  # For reproducibility
  folds <- sample(1:k, nrow(data), replace = TRUE)
  
  # Loop through each fold
  for (i in 1:k) {
    
    # Create training and test sets based on the fold
    test_indices <- which(folds == i)
    train_indices <- setdiff(1:nrow(data), test_indices)
    
    # Create training and test datasets
    train_data <- data[train_indices, ]
    test_data <- data[test_indices, ]
    
    # Fit the model using glmmTMB on the training data
    model <- glmmTMB(formula, data = train_data)
    
    # Predict on the test set
    predictions <- predict(model, newdata = test_data, type = "response")
    
    # Calculate MAE for the test set
    rmse_values[i] <- sqrt(mean((predictions - test_data$covid_mortality_rate_100k)^2))
  }
  
  # Return the average MAE across all folds
  mean_rmse <- mean(rmse_values)
  return(mean_rmse)
}


# Run 10-fold cross-validation for the multilevel model
cv_result <- k_fold_cv_glmmTMB(df_mort_95, k = 20, formula = model_formula)

# View the average RMSE result
print(cv_result) #32.5

127 / (127 + 1013) #0.1114035

anova(glmm2, glmm4, glmm5)

compare_performance(glmm1, glmm2, glmm3, glmm4, glmm5)
```


#MAE 

```{r}
#Predictions and MAE, boostrapping 95% CI

predictions1 <- predict(glmm1, newdata = df_mort_95)
predictions2 <- predict(glmm2, newdata = df_mort_95)
predictions3 <- predict(glmm3, newdata = df_mort_95)
predictions4 <- predict(glmm4, newdata = df_mort_95)
predictions5 <- predict(glmm5, newdata = df_mort_95)

actual_values <- df_mort_95$covid_mortality_rate_100k

mae1 <- mean(abs(actual_values - predictions1))
mae2 <- mean(abs(actual_values - predictions2))
mae3 <- mean(abs(actual_values - predictions3))
mae4 <- mean(abs(actual_values - predictions4))
mae5 <- mean(abs(actual_values - predictions5))

# Example Data
# Assuming df_mort_95 is your dataset and glmm1, glmm2, glmm3, glmm4 are your models

# Fit your models (assuming these are already fitted)
# glmm1 <- glmer(..., data = df_mort_95)
# glmm2 <- glmer(..., data = df_mort_95)
# glmm3 <- glmer(..., data = df_mort_95)
# glmm4 <- glmer(..., data = df_mort_95)

# Function to calculate MAE for a bootstrap sample
calculate_mae <- function(data, model) {
  # Resample the data with replacement
  bootstrap_sample <- data[sample(nrow(data), replace = TRUE), ]
  
  # Generate predictions for the bootstrap sample
  bootstrap_predictions <- predict(model, newdata = bootstrap_sample, re.form = NULL)
  
  # Calculate and return the MAE for the bootstrap sample
  mae <- mean(abs(bootstrap_sample$covid_mortality_rate_100k - bootstrap_predictions))  # Replace 'Y' with the actual dependent variable name
  return(mae)
}

# Number of bootstrap samples (e.g., 1000 iterations)
n_bootstrap <- 1000

# Function to calculate CI for MAE of a model
calculate_ci_for_mae <- function(model, data, n_bootstrap = 1000) {
  mae_values <- numeric(n_bootstrap)
  
  # Perform bootstrap resampling and calculate MAE for each sample
  for (i in 1:n_bootstrap) {
    mae_values[i] <- calculate_mae(data, model)
  }
  
  # Calculate the 95% confidence interval for MAE
  ci_lower <- quantile(mae_values, 0.025)
  ci_upper <- quantile(mae_values, 0.975)
  
  return(c(ci_lower, ci_upper))
}

# Calculate confidence intervals for each model
ci_glmm1 <- calculate_ci_for_mae(glmm1, df_mort_95, n_bootstrap)
ci_glmm2 <- calculate_ci_for_mae(glmm2, df_mort_95, n_bootstrap)
ci_glmm3 <- calculate_ci_for_mae(glmm3, df_mort_95, n_bootstrap)
ci_glmm4 <- calculate_ci_for_mae(glmm4, df_mort_95, n_bootstrap)
ci_glmm5 <- calculate_ci_for_mae(glmm5, df_mort_95, n_bootstrap)

# Print the confidence intervals for each model
cat("95% CI for MAE (GLMM 1): [", ci_glmm1[1], ", ", mae1, ", ", ci_glmm1[2], "]\n")
cat("95% CI for MAE (GLMM 2): [", ci_glmm2[1], ", ", mae2, ", ", ci_glmm2[2], "]\n")
cat("95% CI for MAE (GLMM 3): [", ci_glmm3[1], ", ", mae3, ", ", ci_glmm3[2], "]\n")
cat("95% CI for MAE (GLMM 4): [", ci_glmm4[1], ", ", mae4, ", ", ci_glmm4[2], "]\n")
cat("95% CI for MAE (GLMM 4): [", ci_glmm5[1], ", ", mae5, ", ", ci_glmm5[2], "]\n")
AIC(glmm1)
AIC(glmm2)
AIC(glmm3)
AIC(glmm4)

simulationOutput <- simulateResiduals(fittedModel = glmm5, plot = F)
plot(simulationOutput)

```



#Predictions by strata for each model

```{r}

models <- list(glmm1, glmm2, glmm3, glmm4, glmm5)

# Define strata of interest (example: subset based on a certain variable, e.g., "group")
strata_data1 <- subset(df_mort_95, strata == "Maj_Male , 1_age , below median , low median income")
strata_data1$strata <- "Majority Male, Q1 median age, below median %ile non-white population, Q1 median income (n = 11)"
strata_data2 <- subset(df_mort_95, strata == "Maj_Male , 4_age , above median , high median income")
strata_data2$strata <- "Majority Male, Q4 median age, above median %ile non-white population, Q4 median income (n = 12)"
strata_data3 <- subset(df_mort_95, strata == "Maj_Female , 2_age , above median , high median income")
strata_data3$strata <- "Majority Female, Q2 median age, above median %ile non-white population, Q4 median income (n = 107)"
strata_data4 <- subset(df_mort_95, strata == "Maj_Male , 3_age , below median , med-high median income")
strata_data4$strata <- "Majority Male, Q3 median age, below median %ile non-white population, Q3 median income (n = 100)"

strata_data <- rbind(strata_data1, strata_data2, strata_data3, strata_data4)

# Create a function to extract predicted values for a specific strata
get_predictions <- function(model, newdata) {
  predict(model, newdata = newdata, type = "response", re.form = NA)  # or use re.form = NULL depending on random effects
}

# Extract predictions for each model and strata
predictions_list <- lapply(models, get_predictions, newdata = strata_data)

# Combine predictions into a single data frame for comparison
predictions_df <- data.frame(strata_data, do.call(cbind, predictions_list))

# Rename the columns for clarity
colnames(predictions_df)[(ncol(strata_data)+1):ncol(predictions_df)] <- paste("Model", 1:5)


# View the resulting dataframe with predictions for each model
head(predictions_df)

predictions_long <- predictions_df %>%
  gather(key = "Model", value = "Predicted", `Model 1`:`Model 5`) %>%
  select(strata, Predicted, covid_mortality_rate_100k, Model)

real_data <- select(strata_data, covid_mortality_rate_100k)
real_data$Model <- "Observed"
real_data$strata <- strata_data$strata
real_data$Predicted <- real_data$covid_mortality_rate_100k

predictions_long <- rbind(predictions_long, real_data) 

by(predictions_long$Predicted, predictions_long$Model, summary)

#mean values

mean_values <- predictions_long %>%
  group_by(Model, strata) %>%
  summarise(Mean_Pred = mean(Predicted, na.rm = TRUE))

mean_observed <- predictions_long %>%
  filter(Model == "Observed") %>%
  summarise(Mean_Pred = mean(Predicted, na.rm = TRUE))


observed_means <- predictions_long %>%
  filter(Model == "Observed") %>%
  group_by(strata) %>%
  summarise(Observed_Mean = mean(Predicted, na.rm = TRUE))


global_mean <- mean(df_mort_95$covid_mortality_rate_100k)

ggplot(predictions_long, aes(x = Model, y = Predicted, group = Model, color = Model)) +
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = "point", shape = 20, size = 2, color = "black", fill = "black") +
  labs(
    title = "Figure 1: Comparison across models for strata of low and high membership",
    subtitle = "With global mean (dotted line) and strata mean (solid line)",
    x = "Models",
    y = "Predicted Values"
  ) +
  theme_minimal() +
  facet_wrap(~strata) +  # Facet by strata
  # Add a horizontal line at the mean of "observed" for each strata
  geom_hline(data = observed_means, aes(yintercept = Observed_Mean, color = "Observed Mean"),
             linetype = "solid", size = 0.4, color = "black") +
  geom_hline(aes(yintercept = global_mean), linetype = "dashed", size = 0.2, color = "black")


```





#Bayesian approach
This is not included in the manuscript but it provides an alternative approach closely aligned with previous examples of MAIHDA.
For ease of comparison, GLMM was used in the manuscript, but the BRMS models produce the same results. 

```{r}

brms1 <- brm(covid_mortality_rate_100k ~ 1 + (1 | strata), df_mort_95, family = "gaussian",
                            warmup = 5000, iter = 10000, 
                            chains = 1, seed = 195)
summary(brms1)
loo(brms1) #looic 30595.5
mcmc_plot(brms1)
icc(brms1) #0.352
plot(brms1, variable = "^b", regex = TRUE)

#VPC 
ranef_var <- VarCorr(brms1)

strata_sd <- ranef_var$strata[[1]]
strata_var <- strata_sd^2

residual_sd <- ranef_var$residual__[[1]]
residual_var <- residual_sd^2

vpc_strata <- strata_var / (strata_var + residual_var)
vpc_strata #0.3455358 (0.2791841, 0.4141902)

#Model 2

brms2 <- brm(covid_mortality_rate_100k ~ standardise(median_age_total) + standardise(median_income_12_mo) + standardise(P_county_male_pop) + standardise(P_non_white_only_pop) + (1 | strata_sex_cat:strata_age_cat:strata_income_cat:strata_race_cat), df_mort_95, family = "gaussian")
summary(brms2)
loo(brms2) #looic 30463.3
mcmc_plot(brms2)
icc(brms2) #0.134

#VPC 
ranef_var <- VarCorr(brms2)

strata_sd <- ranef_var$strata[[1]]
strata_var <- strata_sd^2

residual_sd <- ranef_var$residual__[[1]]
residual_var <- residual_sd^2

vpc_strata <- strata_var / (strata_var + residual_var)
vpc_strata #0.1300675 (0.08964915, 0.1821352)

#PCV

0.1197096 / 0.3472363 #0.3447497

conditional_effects(brms2)
loo(brms2) #looic 30479.8
pp_check(brms2)


#Model 3 traditional MAIHDA

brms3 <- brm(covid_mortality_rate_100k ~ strata_sex_cat + strata_age_cat + strata_income_cat + strata_race_cat + (1 | strata_sex_cat:strata_age_cat:strata_income_cat:strata_race_cat), df_mort_95, family = "gaussian")
summary(brms3)
conditional_effects(brms3)
loo(brms3) #looic 30587.2
pp_check(brms3)

#VPC 
ranef_var <- VarCorr(brms3)

strata_sd <- ranef_var$strata[[1]]
strata_var <- strata_sd^2

residual_sd <- ranef_var$residual__[[1]]
residual_var <- residual_sd^2

vpc_strata <- strata_var / (strata_var + residual_var)
vpc_strata #0.04939537 (0.02700924, 0.0810209)

#PCV

0.04939537 / 0.3455358 #0.142953


#model 4

brms4 <- brm(covid_mortality_rate_100k ~ strata_sex_cat:strata_age_cat:strata_income_cat:strata_race_cat, df_mort_95, family = "gaussian")
summary(brms4)


#model 5

brms5 <- brm(covid_mortality_rate_100k ~ standardise(median_age_total)*standardise(median_income_12_mo)*standardise(P_county_male_pop)*standardise(P_non_white_only_pop) + (1 | strata_sex_cat:strata_age_cat:strata_income_cat:strata_race_cat), df_mort_95, family = "gaussian")
summary(brms5)
plot(brms5)
conditional_effects(brms5)
loo(brms2) #looic 30479.8
pp_check(brms2)

#VPC 
ranef_var <- VarCorr(brms5)

strata_sd <- ranef_var$strata[[1]]
strata_var <- strata_sd^2

residual_sd <- ranef_var$residual__[[1]]
residual_var <- residual_sd^2

vpc_strata <- strata_var / (strata_var + residual_var)
vpc_strata #0.1212295. (0.08, 0.18)


#comparison

loo(brms1, brms2, brms3, brms5)

```

#Bayesian

```{r}
brms1 <- brm(covid_mortality_rate_100k ~ 1 + (1 | strata_sex_cat:strata_age_cat:strata_income_cat:strata_race_cat), df_mort_95, family = "gaussian")
summary(brms1)
check_model(brms1)
loo(brms1) #looic 30676


brms2 <- brm(covid_mortality_rate_100k ~ standardise(median_age_total) + standardise(median_income_12_mo) + standardise(P_county_male_pop) + standardise(P_non_white_only_pop) + (1 | strata_sex_cat:strata_age_cat:strata_income_cat:strata_race_cat), df_mort_95, family = "gaussian")
summary(brms2)
plot(brms2)


brms3 <- brm(covid_mortality_rate_100k ~ strata_sex_cat + strata_age_cat + strata_income_cat + strata_race_cat + (1 | strata_sex_cat:strata_age_cat:strata_income_cat:strata_race_cat), df_mort_95, family = "gaussian")
summary(brms3)


loo(brms1, brms3)

brms4 <- brm(covid_mortality_rate_100k ~ strata_sex_cat:strata_age_cat:strata_income_cat:strata_race_cat, df_mort_95, family = "gaussian")
summary(brms4)


brms5 <- brm(covid_mortality_rate_100k ~ standardise(median_age_total)*standardise(median_income_12_mo)*standardise(P_county_male_pop)*standardise(white_non_ratio) + (1 | strata_sex_cat:strata_age_cat:strata_income_cat:strata_race_cat), df_mort_95, family = "gaussian")
summary(brms5)
plot(brms5)


```


